---
title: "Untitled"
author: "Prachi Sardana"
date: "2024-03-23"
output: html_document
---


Loading the required libraries 

```{r}

library(DBI)
library(RSQLite)

```


Step1 - Identifying the dimensions of each dimension table 

```{r}


# Reading the first dimension table business.csv file 
business_csv_data <- read.csv("C:/Users/sarda/Downloads/business.csv")

# Displaying the dimensions of business csv data 
dim(business_csv_data)


```


```{r}


# Displaying the first 10 rows of business csv data 
head(business_csv_data, 10)



```


```{r}

# Reading the second dimension table bed_type 
bed_type_csv <- read.csv("C:/Users/sarda/Downloads/bed_type.csv")

# Displaying the dimension of bed_type csv data 
dim(bed_type_csv)



```


```{r}

# Displaying the first 10 rows of bed_type csv data 
head(bed_type_csv, 10)

```


```{r}

# Reading the fact table fact_csv 
bed_fact_csv <- read.csv("C:/Users/sarda/Downloads/bed_fact.csv")

# Displaying the dimensions of bed_fact_csv data 
dim(bed_fact_csv)

```

```{r}


head(bed_fact_csv, 10)


```

In business csv data , there are 6 vaiables "ims_org_id," "business_name," "ttl_license," "ttl_census," "ttl_staffed," and "bed_cluster_id" and out of these variables, ‘ttl_license,’ ‘ttl_census,’ and ‘ttl_staffed’  are quantitative measures i.e these are numerical values which can change over time and these can be considered as facts since these variables provide information about the licensing , census and staffing levels of various entities. 
While ims_org_id serves as a primary key which identifying each organization and the bed_cluster_id categorize beds into different clusters and each cluster represents a distinct category or group of beds , business name a non measurable entity and hence these 3 variables serve as dimensions. 

In bed_type csv table, there are 3 variables including 'bed_id', 'bed_code', 'bed_desc'are dimension variables which remain constant and don't change over time. These variables provide descriptive information related to bed types with their code and id. 

In bed_fact_csv table ,  ‘ims_org_id,’ ‘business_name,’ and ‘bed_cluster_id are three variables which are non measurable and hence these variables serve as dimensions which tells descriptive information, consist of unique identifiers and categorical groupings that do not represent the quantifiable data. 

Overall business.csv contain the measurable facts and descriptive dimensions while bed_type csv consist of only descriptive dimensions. 



Step 2: Identify the Facts variables from the single Fact Table

In business csv data , there are 6 vaiables "ims_org_id," "business_name," "ttl_license," "ttl_census," "ttl_staffed," and "bed_cluster_id" and out of these variables, ‘ttl_license,’ ‘ttl_census,’ and ‘ttl_staffed’  are quantitative measures i.e these are numerical values which can change over time and these can be considered as facts since these variables provide information about the licensing , census and staffing levels of various entities. 

In bed_fact.csv, the 'license_beds', 'census_beds', and staffed_beds' serve as facts since these are the measurable entities and can change over time. These variables are numerical values that provide information about the licensing, census, and staffing levels, making them suitable for analysis and tracking.



Step 3a: Analysis for Leadership 


```{r}

# Specifying the path of creating_database 
creating_database <-  "C:/Project Machine learning/DA5020/practicum2.db"


# Creating a database connection to connect the databse to creating_database variable
db_connection <- dbConnect(RSQLite::SQLite(), dbname = creating_database)


# Writing the business csv data to SQlite database  where the name specifies the data to be written in the table and value represents the name of table in database.

dbWriteTable(db_connection, name = "business_csv_data", value = business_csv_data, overwrite = TRUE)

# Writing the bed type csv data to SQlite database where the name specifies the data to be written in the table and value represents the name of table in database.

dbWriteTable(db_connection, name = "bed_type_csv", value = bed_type_csv, overwrite = TRUE)


# Writing the bed_fact csv data to SQlite database where the name specifies the data to be written in the table and value represents the name of table in database.
dbWriteTable(db_connection, name = "bed_fact_csv", value = bed_fact_csv, overwrite = TRUE)


```

Step3a 

Identify which hospitals have an Intensive Care Unit (ICU bed_id = 4) bed or a Surgical Intensive Care
Unit (SICU bed_id = 15) bed or both
Create three summary reports that show the following:

1. License beds: List of Top 10 Hospitals ordered descending by the total ICU or SICU license beds.

Include just two variables, hospital_name (business_name) and the total license beds from above as one summary fact. But include only 10 rows in your output table.



```{r}




License_beds <- dbGetQuery(db_connection, "SELECT DISTINCT business_csv_data.business_name AS hospital_name,
                            SUM(CASE WHEN bed_fact_csv.bed_id = 4 OR bed_fact_csv.bed_id = 15 THEN bed_fact_csv.license_beds ELSE 0 END) AS total_license_beds
                        FROM business_csv_data
                        INNER JOIN bed_fact_csv ON business_csv_data.ims_org_id = bed_fact_csv.ims_org_id
                        GROUP BY hospital_name
                        ORDER BY total_license_beds DESC
                        LIMIT 10;")


print(License_beds)




```

2) Do the same thing for Census beds. List of Top 10 Hospitals ordered by total icu or sicu census beds.
Include just two variables, hospital_name (business_name) and the total census beds from above as one
summary fact. Include only 10 rows again.




```{r}


Census_beds <- dbGetQuery(db_connection, "SELECT DISTINCT business_csv_data.business_name AS hospital_name,
                            SUM(CASE WHEN bed_fact_csv.bed_id = 4 OR bed_fact_csv.bed_id = 15 THEN bed_fact_csv.census_beds ELSE 0 END) AS total_census_beds
                        FROM business_csv_data
                        INNER JOIN bed_fact_csv ON business_csv_data.ims_org_id = bed_fact_csv.ims_org_id
                        GROUP BY hospital_name
                        ORDER BY total_census_beds DESC
                        LIMIT 10;")


print(Census_beds)



```


3) Do the same thing for Staffed beds. List of Top 10 Hospitals ordered by the total icu or sicu staffed
beds. Include just two variables, hospital_name (business_name) and the sum of staffed beds from
above as one summary fact. Include only 10 rows again

```{r}



Staffed_beds <- dbGetQuery(db_connection, "SELECT DISTINCT business_csv_data.business_name AS hospital_name,
                            SUM(CASE WHEN bed_fact_csv.bed_id = 4 OR bed_fact_csv.bed_id = 15 THEN bed_fact_csv.staffed_beds ELSE 0 END) AS total_staffed_beds
                        FROM business_csv_data
                        INNER JOIN bed_fact_csv ON business_csv_data.ims_org_id = bed_fact_csv.ims_org_id
                        GROUP BY hospital_name
                        ORDER BY total_staffed_beds DESC
                        LIMIT 10;")


print(Staffed_beds)





```




