---
title: "Practice / Getting to know R"
author: "Prachi Sardana"
date: "15/09/23"
output:
  html_document:
    df_print: paged
---

## Question1 

To load data csv file into a dataframe df.flights and used stringAsFactors = False so that the character coloumns don't convert into the factors.

```{r}
df.flights <- read.csv(url("https://s3.us-east-2.amazonaws.com/artificium.us/datasets/FlightsWithAirlines.csv"),stringsAsFactors = FALSE)

```

## Question 2 

The structure of data consitute 18 observation of 14 Variables showing the summary of dataframe

```{r Q2_str , echo= FALSE}

str(df.flights)

```

## Question 3

used the head function to display first 4 rows

```{r Q3_head , echo = FALSE}

head(df.flights,4)

```

## Question 4

used tail function to display last 5 rows of a data frame

```{r Q4_tail ,echo=FALSE}

tail(df.flights,5)

```

## Question 5

selected carrier, flight, origin, and destination columns using a vector from the data frame

```{r Q5_}

specific_col <- df.flights[,c("carrier","flight","origin","dest")]

print(specific_col)

```


## Question 6 

To calculate the average mean departure delay and display the result in R using cat function

```{r Q6_}

mean_dep_delay <- mean(df.flights$dep_delay)

cat("Average(mean) departure delay is",round(mean_dep_delay),1)

```

## Question 7 

Created an empty column tod to the data frame 

used if else statement based on the syntax ifelse(test_expression,x,y)

If the departure hour is less than 10 , fill the value of am otherwise pm 

Displayed the carrier, flight, dep_hr, and the new tod columns.


```{r Q7_}
# Created an empty dataframe tod

df.flights$tod <- 0

# used ifelse statement based on the syntax ifelse(test_expression, x, y)

# where test_expression is the logical condition or vector, x is the value to return when condition = True and y is the value to return when condition = False

df.flights$tod <- ifelse(df.flights$dep_hr < 12, "am", "pm")

specific_col <- df.flights[,c("carrier","flight","dep_hr","tod")]
print(specific_col)

```


## Question 8 

For each flight, display the carrier, flight number, and the actual departure time (scheduled departure plus departure delay) for flights that were delayed. Display the time in the format hh:mm in 24-hour format, e.g., display 23:20 rather than 11:20 or 11:20PM.

```{r}

# creating a data frame delayed_flights by subsetting the df.flights that have departure delay greater than 0 and includes the following coloumns

delayed_flights <- df.flights[df.flights$dep_delay > 0,c("carrier", "flight","dep_hr","dep_min","dep_delay")]


# actual departure = scheduled departure time in min + departure hour * 60 + the departure delay in min 

actual_departure <- delayed_flights$dep_min + delayed_flights$dep_hr*60 + delayed_flights$dep_delay

# create a new coloumn act_dep_hh in the delayed flights  
# as.POSIXlt function covert the time into list of date-time components
# format function is used to extract and format the "hour:minute" component of the date-time in a 24-hour format ("%H:%M").


delayed_flights$act_dep_hh <- format(as.POSIXlt(actual_departure, origin = "2023-09-01"), format = "%H:%M")
print(delayed_flights)


```

