---
title: "Working with Data Frames in R"
author: "Prachi Sardana"
output:
  word_document: default
  html_notebook: default
---

## Question 1
 
To load data csv file into a data frame df.wines

```{r Q1_loadCSV}
df.wines <- read.csv(url("https://s3.us-east-2.amazonaws.com/artificium.us/datasets/whitewines.csv"),header = TRUE,stringsAsFactors = FALSE)

```

## Question 2

Inspect the data frame usig head function and determine it's structure

```{r Q2_ echo = FALSE}

head(df.wines,5)

str(df.wines)
```
## Question 3

How many wines have a residual sugar level above 0.5 and an alcohol level above 7?

```{r Q3_ }

# Subseting df.wines  data frame to determine how many wines have residual sugal above 0.5 and alcohol level above 7

filter_df.wines <- subset(df.wines, df.wines$residual.sugar > 0.5 & df.wines$alcohol > 7)

# To count the number of rows 

count_rows <- nrow(filter_df.wines)

# Using cat function to print the count of wines

cat("Number of wines with Residual Sugar > 0.5 and Alcohol > 7:", count_rows, "\n")



```
## Question 4 

How many wines have an alcohol content between 9.5 and 11.5 (exclusive) and a quality rating below 7?

```{r Q4_ }
# used subset function to filter out alcohol content between 9.5 and 11.5(exclusive) and quality rating above 7 

filter_df.wines <- subset(df.wines, df.wines$alcohol >= 9.5 & df.wines$alcohol <= 11.5 & df.wines$quality < 7)

# Counting rows using nrow from filter_df.wines
count_rows <- nrow(filter_df.wines)

# Using cat function to print the count of wines

cat("Number of wines with alcohol content between 9.5 and 11.5 (exclusive) and a quality rating below 7:", count_rows, "\n")


```
## Question 5

Are there any wines with an alcohol content above 14?

```{r Q5_}
# used subset function to filter df.wines containing alcohol content greater than 14

filter_df.wines <- subset(df.wines, df.wines$alcohol > 14)

# Counting rows using nrow for filter_df.wines

count_rows <- nrow(filter_df.wines)

# Using cat function to print the count of wines
cat("Number of wines with alcohol content above 14:", count_rows, "\n")

```
## Question 6

What are the median and mean alcohol content of all wines?

```{r Q6_ }

# mean alcohol content 

mean_alcohol <- mean(df.wines$alcohol)
print(mean_alcohol)

# median alcohol content 

median_alcohol <- median(df.wines$alcohol)
print(median_alcohol)

```
## Question 7 

Create a new column called *swill_index* in the data frame *df.wines* that is determined by multiplying the residual sugar content by the alcohol content and dividing the alcohol content by the quality.

```{r Q7_}

df.wines$swill_index <- (df.wines$alcohol/df.wines$quality)* df.wines$residual.sugar


```
## Question 8 

The data frame *df.wines* should now have a new column named *alcohol.z* that determines the alcohol's *z*-score. The amount of standard deviations a value is from the mean is used to compute a feature's *z*-score. Create the alcohol.z column with the value: *mean(alcohol) - alcohol] / sd(alcohol)*. Then compute the mean and standard deviation for alcohol.

```{r Q8_}

# Mean alcohol content
mean_alcohol <- mean(df.wines$alcohol)

# Standard deviation 
sd_alcohol <- sd(df.wines$alcohol)

# Calculating the z score and adding the new coloumn to the data frame which is calculated by x-u/sd  formula
df.wines$alcohol.z <- (mean_alcohol - df.wines$alcohol)/sd_alcohol

```

